### 2019-07-25 0/exams/f-practice.tex
Practice exam.

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 2 (Practice)}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2017 \ \\
  }
\end{center}

@@

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 12 \\ \hline
  2 & 12 \\ \hline
  3 & 15 \\ \hline
  4 & 15 \\ \hline
  5 & 15 \\ \hline
  6 & 10 \\ \hline
  7 & 15 \\ \hline \hline
  Total: & 94\\
  \hline
\end{tabular}
\end{center}

\noindent
This test is open-book, open-notes, but you may not use any computing
device other than your brain and may not communicate with anyone.
You have 50 minutes to complete the test.

@@

\begin{verbatim}
Boolean String -> String
[Boolean -> String] -> String
[Listof Boolean] [Listof Number] -> Number
[X] . [Number -> X] -> X
[X Y] . [X -> Y] [Listof X] -> [Listof Y]
[X Y Z] . [X -> Y] [Y -> Z] X -> Z
\end{verbatim}
Define one example of a function in each class.  (You only need to
provide the code, not a full design.)

@@

\noindent
{\bf Problem 2 (12 points).}
%
For each of the following functions, provide the most general signature
that correctly describes the function:

\begin{alltt}
;; f : \censor{Number -> Number}

(define (f x)
  (+ (sqr x) 2))

;; g : \censor{[Listof Number] -> [Listof Number]}

(define (g x)
  (map add1 x))

;; h : \censor{[Listof Number] [Listof Number] -> [Listof Number]}

(define (h x y)
  (cond [(empty? x) y]
        [(cons? x)
         (cons (sqr (first x))
               (h (rest x) y))]))

;; i : \censor{(Number -> Number) -> Number}

(define (i x)
  (sqr (x 5)))

;; j : \censor{Number -> (Number -> Number)}

(define (j x)
  (lambda (y)
    (+ x y)))

;; k : \censor{[X] (Natural -> X) Natural -> [Listof X]}

(define (k x y)
  (cond [(= y 0) (list (x 0))]
        [else
          (cons (x y) (k x (sub1 y)))]))
\end{alltt}

@@

;; avg : [Listof Number] -> Number
;; Compute the average of a list of numbers
;; Assume: the list is non-empty
(check-expect (avg (list 3 4 5)) 4)
(check-expect (avg (list 10 20)) 15)
(define (avg lon)
  (/ (sum lon) 
     (length lon)))

;; sum : [Listof Number] -> Number
;; Sum the list of numbers
(check-expect (sum '()) 0)
(check-expect (sum (list 3 4 5)) 12)
(define (sum lon)
  (foldr + 0 lon))

;; Alt: follow [Listof Number] template
\end{SaveVerbatim}

@@

\newpage 
\noindent
{\bf Problem 4 (15 points).}
%
Here is a parametric data definition for a non-empty list of items:
\begin{verbatim}
;; A [NEListof X] is one of:
;; - (cons X '())
;; - (cons X [NEListof X])
;; Interp: an arbitrarily long but non-empty sequence of items.
\end{verbatim}
Here are two similar programs that operate over non-empty lists:
\begin{verbatim}
;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;

;; smallest : [NEListof Number] -> Number
;; Produce the smallest number in the list
(check-expect (smallest (list 6 3 2 4)) 2)
(define (smallest xs)
  (cond [(empty? (rest xs)) (first xs)]
        [else
          (min (first xs)
               (smallest (rest xs)))]))

@@

;; longest : [NEListof String] -> String
;; Produce the first longest string in the list
(check-expect (longest (list "hi" "there")) "there")
(check-expect (longest (list "same" "size")) "same")
(define (longest xs)
  (cond [(empty? (rest xs)) (first xs)]
        [else
          (longer (first xs)
                  (longest (rest xs)))]))

@@

;; mostest : [X] [NEListof X] [X X -> X] -> X
;; Produce the first most-est element in the list
(define (mostest xs most)
  (cond [(empty? (rest xs)) (first xs)]
        [else
          (most (first xs)
                (mostest (rest xs) most))]))

@@

\newpage
\noindent
{\bf Problem 5 (15 points).}  
%
Design a program that takes a list of numbers and produces the count
of positive numbers in the list.  For example, if the list contains 3,
4, -1, 0, -2, and 5, the count of positive numbers is 3.  You may
assume the {\tt{[Listof Number]}} data definition is defined.
%
For full credit, use list abstraction functions.  For partial credit,
follow the template for {\tt{[Listof Number]}}.

@@

;; count-pos : [Listof Number] -> Number
;; Count the number of non-negative numbers in given list
(check-expect (count-pos '()) 0)
(check-expect (count-pos (list 3 -1 0 2)) 2)
(define (count-pos lon)
  (foldr (lambda (n c) (if (positive? n) (add1 c) c)) 0 lon))

@@

\noindent
{\bf Problem 6 (10 points).}
%
It's been said that {\tt foldr} is \emph{the} fundamental list
abstraction function because any function that follows the list
template can be expressed in terms of {\tt foldr}.  Put that theory to
the test by providing an equivalent definition for the following
function in terms of {\tt foldr}:

\begin{verbatim}
;; map : [X Y] . [X -> Y] [Listof X] -> [Listof Y]
;; Apply f to each element of list, collect a list of results
(define (map f ls)
  (cond [(empty? ls) '()]
        [(cons? ls)
         (cons (f (first ls))
               (map f (rest ls)))]))
\end{verbatim}

\noindent
(Just provide code, not a full design.)

(define (map f xs)
  (foldr (lambda (x ys) (cons (f x) ys)) '() xs))
\end{SaveVerbatim}

\noindent
{\bf Problem 7 (15 points).}
%
A sports tournament in which teams pair off against eachother and
the winner advances to the next level in the tournament can be modelled
with the following data definition:

>>>>

\begin{center}
  {\Large {\bf CMSC 198Q, Final Exam (Practice)}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Summer 2019 \ \\
  }
\end{center}

@@

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 12 \\ \hline
  2 & 15 \\ \hline
  3 & 15 \\ \hline
  4 & 15 \\ \hline
  5 & 15 \\ \hline
  6 & 15 \\ \hline \hline
  Total: & 87\\
  \hline
\end{tabular}
\end{center}

\noindent
This test is open-book, open-notes, but you may not use any computing
device other than your brain and may not communicate with anyone.
You have 120 minutes to complete the test.

@@

\begin{verbatim}
Boolean String -> String
[Boolean -> String] -> String
[Lo Boolean] [Lo Number] -> Number
[Number -> X] -> X
[X -> Y] [Lo X] -> [Lo Y]
[X -> Y] [Y -> Z] X -> Z
\end{verbatim}
Define one example of a function in each class.  (You only need to
provide the code, not a full design.)

@@

\noindent
{\bf Problem 2 (15 points).} Here is a data definition:
\begin{verbatim}
;; A Zoinks is one of:
;; - Jinkies
;; - Jeepers

;; A Jinkies is a (make-velma String Number)
(define-struct velma (my glasses))

;; A Jeepers is a (make-daphne Zoinks Zoinks)
(define-struct daphne (ruh roh))
\end{verbatim}

\noindent
1) Define a template for functions that consume the following kind of data.

\noindent
2) Give three different examples of {\tt Zoinks}, including at least one {\tt Jeepers}.

(define (zoinks-template z)
  (cond [(velma? z) (... (velma-my z) (velma-glasses z) ...)]
        [(daphne? z) 
         (... (zoinks-template (daphne-ruh z))
              (zoinks-template (daphne-roh z)) ...)]))

(make-velma "" 0)
(make-velma "a" 1)
(make-daphne (make-velma "" 0)
             (make-velma "a" 1))

\end{SaveVerbatim}

@@

;; avg : [Lo Number] -> Number
;; Compute the average of a list of numbers
;; Assume: the list is non-empty
(check-expect (avg (list 3 4 5)) 4)
(check-expect (avg (list 10 20)) 15)
(define (avg lon)
  (/ (sum lon) 
     (length lon)))

;; sum : [Lo Number] -> Number
;; Sum the list of numbers
(check-expect (sum '()) 0)
(check-expect (sum (list 3 4 5)) 12)
(define (sum lon)
  (foldr + 0 lon))

;; Alt: follow [Lo Number] template
\end{SaveVerbatim}

@@

\newpage 
\noindent
{\bf Problem 4 (15 points).}
%
Here is a parametric data definition for a non-empty list of items:
\begin{verbatim}
;; A [NELo X] is one of:
;; - (cons X '())
;; - (cons X [NELo X])
;; Interp: an arbitrarily long but non-empty sequence of items.
\end{verbatim}
Here are two similar programs that operate over non-empty lists:
\begin{verbatim}
;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;

;; smallest : [NELo Number] -> Number
;; Produce the smallest number in the list
(check-expect (smallest (list 6 3 2 4)) 2)
(define (smallest xs)
  (cond [(empty? (rest xs)) (first xs)]
        [else
          (min (first xs)
               (smallest (rest xs)))]))

@@

;; longest : [NELo String] -> String
;; Produce the first longest string in the list
(check-expect (longest (list "hi" "there")) "there")
(check-expect (longest (list "same" "size")) "same")
(define (longest xs)
  (cond [(empty? (rest xs)) (first xs)]
        [else
          (longer (first xs)
                  (longest (rest xs)))]))

@@

;; mostest : [NELo X] [X X -> X] -> X
;; Produce the first most-est element in the list
(define (mostest xs most)
  (cond [(empty? (rest xs)) (first xs)]
        [else
          (most (first xs)
                (mostest (rest xs) most))]))

@@

\newpage
\noindent
{\bf Problem 5 (15 points).}  
%
Design a program that takes a list of numbers and produces the count
of positive numbers in the list.  For example, if the list contains 3,
4, -1, 0, -2, and 5, the count of positive numbers is 3.  You may
assume the {\tt{[Listof Number]}} data definition is defined.
%
For full credit, use list abstraction functions.  For partial credit,
follow the template for {\tt{[Lo Number]}}.

@@

;; count-pos : [Lo Number] -> Number
;; Count the number of non-negative numbers in given list
(check-expect (count-pos '()) 0)
(check-expect (count-pos (list 3 -1 0 2)) 2)
(define (count-pos lon)
  (foldr (lambda (n c) (if (positive? n) (add1 c) c)) 0 lon))

@@

\noindent
{\bf Problem 6 (15 points).}
%
A sports tournament in which teams pair off against eachother and
the winner advances to the next level in the tournament can be modelled
with the following data definition:
